---
title: "Reproducible Research (Project 1)"
author: "Vincent Wysocki"
date: "4/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(chron)
```

**Read Data from WD**
```{r cars}
my_data.with.na <- read.csv("activity.csv")
my_data.with.na$steps <- as.numeric((my_data.with.na$steps), na.rm = TRUE)
my_data <- my_data.with.na %>% filter(my_data.with.na$steps != "NA")
```


**Summing steps, calculating the mean and median values** 
```{r }
sum_steps <- my_data %>% group_by(date) %>% summarise(sum_steps = sum(steps))
ggplot(sum_steps, aes(x = sum_steps)) + geom_histogram(bins = 50) + labs(y = "Number of Days", x = "Steps per Day", title = "Histogram of Step Sum Distribution")

mean_steps  <- my_data %>% group_by(date) %>% summarise(mean(steps))
colnames(mean_steps) <- c("date", "steps")
print(mean_steps$steps)


median_steps <- my_data %>% group_by(date) %>% summarise(median(steps))
colnames(median_steps) <- c("date", "steps")
print(median_steps$steps)
```

**Average the steps, preserving the interval group, and ploting a line chart.** 
```{r }
avg_steps <- my_data %>% group_by(interval) %>% summarise(mean(steps))
colnames(avg_steps) <- c("interval", "steps")
plot(x = avg_steps$interval, y = avg_steps$steps, type = "l", xlab = "Interval", ylab = "Average steps", main  = "Line chart of Average Steps per Interval")
index <- which(avg_steps$steps == max(avg_steps$steps))
print(avg_steps[index, 1])
```



**Finding out how many "NA"s and then replacing that index with the average value for that interval** 
```{r }
index2 <- which(is.na(my_data.with.na$steps))
print(length(index2))
imput_data <- my_data.with.na
for (i in index2){

        imput_data[i,1] <- avg_steps[which(avg_steps$interval == my_data.with.na[i,3]), 2]

}

```


**Finding the steps sum per interval and histogram**  
**The result of imputation is a larger distribution for days with high step counts. Additionally, a higher mean for each day, and median values for some days instead of all being zero**
```{r }

imput.sum_steps <- imput_data %>% group_by(date) %>% summarise(sum_steps = sum(steps))
ggplot(imput.sum_steps, aes(x = sum_steps)) + geom_histogram(bins = 50) + labs(y = "Number of Days", x = "Steps per Day", title = "Histogram of Sum Step Distribution per Day")
```


**Finding Mean and Median Values**
```{r }
imput.mean <- imput_data %>% group_by(date) %>% summarise(mean(steps))
colnames(imput.mean) <- c("date", "steps")
```


**Comparing mean values with imputation vs. no imputation**
```{r }
print(imput.mean$steps)
print(mean_steps$steps)
```

```{r }
imput.median <- imput_data %>% group_by(date) %>% summarise(median(steps))
colnames(imput.median) <- c("date", "steps")
```

**Comparing median values with imputation vs. no imputation**
```{r }
print(imput.median$steps)
print(median_steps$steps)

```


**Setting date as factor and "recoding" for aesthetic of facet_grid**
```{r }

imput_data$date   <- as.Date(imput_data$date)
imput_data$factor <- as.character(is.weekend(imput_data$date))
imput_data$factor <- recode(imput_data$factor, "FALSE" = "Weekday", "TRUE" = "Weekend")
```

**Finding the average steps per interval by day of the week and plotting line graph**
```{r }
new_data <- imput_data %>% group_by(factor, interval) %>% summarize(sum_steps = mean(steps))
ggplot(new_data, aes(x = interval, y = sum_steps)) + geom_line() + coord_cartesian(ylim = c(-2,250)) + 
        facet_grid(factor ~ .) + labs(x = "Intervals", y = "Mean Interval Steps", title  = "Mean Steps per Interval by Day of the Week")

```